---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EcoacousticUtilities

<!-- badges: start -->
[![R-CMD-check](https://github.com/wcornwell/ecoacoustic_utilities/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/wcornwell/ecoacoustic_utilities/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://codecov.io/gh/wcornwell/ecoacoustic_utilities/branch/main/graph/badge.svg)](https://app.codecov.io/gh/wcornwell/ecoacoustic_utilities?branch=main)
<!-- badges: end -->

Utilities for downloading and analyzing ecoacoustic data from iNaturalist and the Atlas of Living Australia (ALA).

## Installation

You can install the development version of EcoacousticUtilities from GitHub:

```{r eval=FALSE}
# install.packages("devtools")
#devtools::install_github("wcornwell/ecoacoustic_utilities")
```

## Functions

### iNaturalist Functions

- **`get_inat_sounds()`** - Download sound recordings from iNaturalist
  - **Use case**: Download actual audio files for a specific taxon
  - With `download=FALSE`: Get quick count of total observations with sounds
  - With `download=TRUE`: Download up to N audio files to local directory
  
- **`get_inat_species_summary()`** - Get species-level summary of sound recordings
  - **Use case**: Identify which species have the most recordings in a taxon/region
  - Returns data.frame with counts per species (taxon_id, scientific_name, common_name, n_recordings)
  - Useful for finding well-documented species before downloading

- **`plot_inat_species_summary()`** - Visualize species summary data
  - **Use case**: Create publication-ready plots of species recording counts
  - Horizontal bar chart with customizable colors and labels

### ALA Functions

- **`get_ala_circle_occurrences()`** - Download occurrence records within a circular area

### Utility Functions

- **`biggest_files()`** - Find the largest files in a directory
- **`find_duplicate_wavs()`** - Detect duplicate WAV files by content hash
- **`training_dataset_summary()`** - Comprehensive analysis of audio training datasets

## Example Usage

```{r eval=FALSE}
library(EcoacousticUtilities)

# Find largest files in a directory
big_files <- biggest_files("/path/to/directory", n = 20)

# Find duplicate WAV files
duplicates <- find_duplicate_wavs("/path/to/audio/library")
```

### iNaturalist Workflow

```{r eval=FALSE}
# 1. Quick check: How many sound recordings exist?
n_sounds <- get_inat_sounds(
  "Turnix maculosus",
  place_name = "Australia",
  download = FALSE  # Just get the count
)
print(n_sounds)  # e.g., 450

# 2. Which species have the most recordings?
species_summary <- get_inat_species_summary(
  taxon_name = "Orthoptera",  # All crickets/grasshoppers
  place_name = "Australia",
  min_recordings = 5  # Only species with 5+ recordings
)
print(species_summary)

# 3. Visualize the results
plot_inat_species_summary(species_summary, top_n = 15)

# 4. Download actual audio files for a specific species
get_inat_sounds(
  "Turnix maculosus",
  place_name = "Australia",
  target_n = 100,
  download = TRUE,  # Download files
  quality = "research"
)
```

### Example: Australian Orthoptera

Here's a real example analyzing cricket and grasshopper recordings in Australia:

```{r example-orthoptera}
library(EcoacousticUtilities)
library(ggplot2)

# Get species with at least 5 recordings
orth_summary <- get_inat_species_summary(
  taxon_name = "Orthoptera",
  place_name = "Australia",
  min_recordings = 5
)

# View top species
head(orth_summary, 10)

# Create visualization
plot_inat_species_summary(
  orth_summary,
  top_n = 15,
  title = "Australian Orthoptera: Top Species by Sound Recordings"
)
```

### ALA Workflow

```{r eval=FALSE}
# Get ALA occurrences within a radius
birds <- get_ala_circle_occurrences(
  taxon = "Aves",
  lat = -33.7,
  lon = 151.3,
  radius_km = 10,
  email = "your.email@example.com"
)
```

### Training Dataset Analysis

```{r eval=FALSE}
# Analyze an audio training dataset
summary <- training_dataset_summary("/path/to/audio/dataset")
print(summary)  # Shows class distribution, file types, sizes, durations, outliers
```

## License

MIT © Z Cornwell
